---
title: "Lab2Seanunfinished"
author: "sbfogelb"
date: "February 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
#SETUP
library(ggplot2)
library(tidyverse)
OD_Tract <- read.csv("OD_TEST_TRACT_Sean.csv", fileEncoding = "latin1")

names(OD_Tract) <- c("Tract", "Census Tract", "Total Number of Jobs", "Workers ages 29 or younger", "Workers ages 30-54", "Workers ages 55 or older", "Number of Jobs earning $1250/month or less", "Number of Jobs earning between $1251 and $3333/month", "Number of jobs earning greater than $3333/month", "Number of jobs in Good Producing industry", "Number of jobs in Trade, Transportation, and Utilities", "Number of jobs in All Other Services")
OD_Tract_5 <- filter(OD_Tract, `Total Number of Jobs` >= 177)
OD_Tract2 = as_tibble(OD_Tract_5)
OD_Tract2$Tract = as.character(OD_Tract2$Tract)
OD_Tract2
```

```{r, echo=FALSE}
#Comparison of the top 5 tracts between the three different work sectors and the total number of jobs

ggplot(OD_Tract2, aes(x=Tract,y=`Total Number of Jobs`, fill = `Number of jobs in Good Producing industry`, label = `Total Number of Jobs`)) + geom_bar(stat = "identity", width=0.8) +geom_text(size = 5, vjust = 5, color = "white") + ggtitle("Total Jobs vs. Number of Jobs in Good Producing Industry") + theme_bw()

ggplot(OD_Tract2, aes(x=Tract,y=`Total Number of Jobs`, fill = `Number of jobs in Trade, Transportation, and Utilities`, label = `Total Number of Jobs`)) + geom_bar(stat = "identity", width=0.8) + geom_text(size = 5, vjust = 5, color = "white") + ggtitle("Total Jobs vs. Number of Jobs in Trade, Transportation, and Utilities Industry") + theme_bw()

ggplot(OD_Tract2, aes(x=Tract,y=`Total Number of Jobs`, fill = `Number of jobs in All Other Services`, label = `Total Number of Jobs`)) + geom_bar(stat = "identity", width=0.8) + geom_text(size = 5, vjust = 5, color = "white") + ggtitle("Total Jobs vs. Number of Jobs in All Other Services industry") + theme_bw()
```
Each graph above compares the total number of jobs in the highest job-populated tracts of Oklahoma to the three different industries where jobs are available. Tracts 1 through 5 correspond to these 5 tracts in order: 3010.07, 1086.01, 3010.08, 9.03. The total number of jobs is a sum of the number of jobs in each industry. If we compare the proportion of jobs from the industries to the total number of jobs, we can see that there is not an even distrubition between the industries in each tract. This could be due to the access of those jobs based on geographic location, as well as social and economic factors. For example, in tract 3, there are over 100 Trade, Transportation, and Utilities jobs, and over 50 Good producing jobs. But there is under 50 jobs disperesed between all other services.